---
title: "SEM Model"
output: pdf_document
date: "`r format(Sys.time(), '%d %B %Y')`"
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(lavaan)
library(semPlot)

opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

## Functions for converting responses to numbers
```{r conversion functions}
# for (course) Grade
GradePoint <- function(x) {
  case_when(
    x == "A" ~ 4.0,
    x == "A-" ~ 3.7,
    x == "B+" ~ 3.3,
    x == "B" ~ 3.0,
    x == "B-" ~ 2.7,
    x == "C+" ~ 2.3,
    x == "C" ~ 2.0,
    x == "C-" ~ 1.7,
    x == "D+" ~ 1.3,
    x == "D" ~ 1.0,
    x == "F" ~ 0.0,
    TRUE ~ NA_real_
  )
}

# for Q18, Q19
TextRead <- function(x) {
  case_when(
    x == "0 to 1" ~ 0.5,
    x == "2 to 3" ~ 2.5,
    x == "4 to 5" ~ 4.5,
    x == "6 to 7" ~ 6.5,
    x == "8 to 9" ~ 8.5,
    x == "10 to 11" ~ 10.5,
    x == "12" ~ 12,
    TRUE ~ NA_real_
  )
}

# for Q20
RJTime <- function(x) {
  case_when(
    x == "less than 1" ~ 0.5,
    x == "1 to <3" ~ 2,
    x == "3 to <6" ~ 4.5,
    x == "6 to <10" ~ 8,
    x == "10 to <15" ~ 12.5,
    x == "more than 15" ~ 18,
    TRUE ~ NA_real_
  )
}

# for Q21
TextEach <- function(x) {
  case_when(
    x == "less than 1" ~ 0.5,
    x == "1 to <2" ~ 1.5,
    x == "2 to <3" ~ 2.5,
    x == "3 to <4" ~ 3.5,
    x == "4 to <5" ~ 4.5,
    x == "more than 5" ~ 5.5,
    TRUE ~ NA_real_
  )
}

# for Q22, Q23
Attend <- function(x) {
  case_when(
    x == "0%" ~ 0,
    x == "1-20%" ~ 0.1,
    x == "21-40%" ~ 0.3,
    x == "41-60%" ~ 0.5,
    x == "61-80%" ~ 0.7,
    x == "81-100%" ~ 0.9,
    TRUE ~ NA_real_
  )
}

EngGood <- function(x) {
  case_when(
    is.na(x) ~ NA_real_,
    x %in% c("DSE", "Others") ~ NA_real_,

    x %in% c(
      "DSE 5 or above",
      "JiangSu Exam >=105",
      "National Exam >=140",
      "Taiwan Exam >=15"
    ) ~ 1,
    TRUE ~ 0
  )
}
```

## Read Data
After reading data, extract useful columns and renaming them.
```{r read data, message=FALSE}
rawdata <- read_csv("RawData/201617MasterData 2018-03-07.csv",
  na = c("BLANK", "NA", "MULT", "#VALUE!", "(4,6)", "", "ERROR #3100", "#N/A")
)

uge <- rawdata %>%
  mutate(
    # background
    EngGood = EngGood(`English Proficiency`),
    FacSci = Faculty %in% c("BASCI", "ENF", "ENSCF", "SCF", "MED"),
    # Effort
    TextRead = TextRead(`Q18 (Assigned Text Read Completely)`),
    RJTime = RJTime(`Q20 (Time/RJ)`),
    TextTimeEach = TextEach(`Q21 (Time/reading)`),
    Lecture = 13 * Attend(`Q22 (% Lecture)`),
    Tutorial = 13 * Attend(`Q23 (% tutorial participation)`),
    ReadTime = TextRead * TextTimeEach,
    # Grade
    GradePoint = GradePoint(Grade)
  ) %>%
  select(
    Sex,
    SciSubj = `ScienceSubject (Phy,Chem,Bio,InterSci,CombSci)`,
    EngGood,
    Faculty,
    FacSci,
    matches("Before|After"),
    matches("Change"),
    TextRead:ReadTime,
    GradePoint
  )

colnames(uge) <- str_replace(colnames(uge), "(Q\\d*|cGPA)\\s*\\(\\s*Before\\s*\\)", "\\1Entry")
colnames(uge) <- str_replace(colnames(uge), "(Q\\d*|cGPA)\\s*\\(\\s*After\\s*\\)", "\\1Exit")
colnames(uge) <- str_replace(colnames(uge), "(Q\\d*)\\s*\\(\\s*Change\\s*\\)", "\\1Change")

```

## Find backgrounds associated with factors
Factors are grouped according to previous years' finding via factor analysis. Backgrounds are chosen based on their correlations with factors before putting them into the model.
```{r find association, warning=FALSE}
uge %>% with({
  f1 <- (Q1Exit + Q3Exit + Q4Exit + Q5Exit) / 4
  f2 <- (Q6Exit + Q7Exit + Q8Exit) / 3
  f3 <- (Q9Exit + Q10Exit + Q11Exit + Q12Exit) / 4
  f4 <- (Q13Exit + Q15Exit) / 2
  f5 <- (Q2Exit + Q14Exit + Q16Exit + Q17Exit) / 4

  SexM <- recode(Sex, "F" = 0, "M" = 1)


  # boxplot of factor vs sex
  print(
    tibble(f1, f2, f3, f4, f5, Sex) %>%
      filter(!is.na(Sex)) %>% # 3 NA case only
      gather("factor", "value", -Sex) %>%
      ggplot(aes(Sex, value)) +
      geom_boxplot() +
      facet_grid(. ~ factor)
  )

  # boxplot of factor vs English Pro
  print(
    tibble(f1, f2, f3, f4, f5, EngGood = factor(EngGood)) %>%
      gather("factor", "value", -EngGood) %>%
      ggplot(aes(EngGood, value)) +
      geom_boxplot() +
      facet_grid(. ~ factor)
  )

  # boxplot of factor vs Science related faculty
  print(
    tibble(f1, f2, f3, f4, f5, FacSci) %>%
      gather("factor", "value", -FacSci) %>%
      ggplot(aes(FacSci, value)) +
      geom_boxplot() +
      facet_grid(. ~ factor)
  )


  tibble(f1, f2, f3, f4, f5, GradePoint, SciSubj, SexM, EngGood, FacSci) %>%
    cor(use = "pair") %>%
    # only show cor btw factors <-> Grade,backgrounds
    .[1:6, -(1:5)]
})
```

## The SEM Model
### Specify the model
```{r model}
sem.model <- '
  # latent variables
    F1Exit =~ Q1Exit + Q3Exit + Q4Exit + Q5Exit
    F2Exit =~ Q6Exit + Q7Exit + Q8Exit
    F3Exit =~ Q9Exit + Q10Exit + Q11Exit + Q12Exit
    F4Exit =~ Q13Exit + Q15Exit
    F5Exit =~ Q2Exit + Q14Exit + Q16Exit + Q17Exit
    
    Effort =~ ReadTime + RJTime + Lecture + Tutorial
    
    Academic =~ GradePoint

  # regressions (outcome ~ background + effort + other latent var)
    Academic ~ EngGood + Effort + F1Exit + F5Exit
    F1Exit ~ EngGood + Effort
    F2Exit ~ SciSubj + FacSci + Effort
    F3Exit ~ SciSubj + FacSci + Effort + F2Exit
    F4Exit ~ Effort + F5Exit
    F5Exit ~ Effort
'
```

### Fit model
The model is fitted and accessed based on CFI, where a value of 0.9 indicates acceptable fit.
```{r fit model}
sem.fit <- sem(sem.model, data = uge, std.lv = TRUE, missing = "fiml")
(cfi <- fitMeasures(sem.fit, fit.measures = "CFI")) # >.9 okay fit
```
In this case, the value of CFI is `r round(cfi, 2)`, which is close to 0.9.

```{r other summary, include=FALSE}
# inspect(sem.fit, 'cor.lv')
# varTable(sem.fit)
# summary(sem.fit, fit.measures=TRUE, standardized=TRUE)
# parameterEstimates(sem.fit, standardized=TRUE)
```

### Factor loading Table
```{r factor loading table}
# nicer table, show composition of factors
parameterEstimates(sem.fit, standardized = TRUE) %>%
  filter(op %in% "=~") %>%
  select("Latent Factor" = lhs, Indicator = rhs, B = est, SE = se, 
         Z = z, "p-value" = pvalue, Beta = std.all) %>%
  kable(digits = 3, format = "pandoc", caption = "Factor Loadings")
```

### Regression Table
```{r Regression Table}
# nicer table, show regression of latent variables on measurements/other latent variables
parameterEstimates(sem.fit, standardized = TRUE) %>%
  filter(op %in% "~") %>%
  select("Latent Factor" = lhs, Predictor = rhs, B = est, SE = se, 
         Z = z, "p-value" = pvalue, Beta = std.all) %>%
  kable(digits = 3, format = "pandoc", caption = "Regression")
```

### Plot Model
```{r plot model}
semPaths(sem.fit, rotation=2,
         residuals = FALSE, exoCov = FALSE, exoVar = FALSE, intercepts = FALSE,
         nCharNodes = 0, font=4, label.cex=2, asize=1.5,
         sizeLat = 5, sizeMan = 8, sizeMan2 =3)
```
For a better image, see the file `semplot.png`.
